---
title: "Project1"
author: "YoungKoung Kim"
date: "February 24, 2018"
output: html_document
---
For Project 1, I used the following packages:
```{r setup, include=FALSE, result = 'asis'}
library(knitr)

```

```{r}
library(stringr)
library(sqldf)
library(tidyr)
```

### Read file by lines
```{r}
file <- readLines("H:/DataScience/CUNY/Spring2018/DATA607/project1/tournamentinfo.txt", n = 195)
```

### Strategy to create a .csv file
The Players information in the chess tournament table is presented in two rows. Each player were sperated by dash lines. 

Here is an excerpt from the file:

```
-----------------------------------------------------------------------------------------
 Pair | Player Name                     |Total|Round|Round|Round|Round|Round|Round|Round| 
 Num  | USCF ID / Rtg (Pre->Post)       | Pts |  1  |  2  |  3  |  4  |  5  |  6  |  7  | 
-----------------------------------------------------------------------------------------
    1 | GARY HUA                        |6.0  |W  39|W  21|W  18|W  14|W   7|D  12|D   4|
   ON | 15445895 / R: 1794   ->1817     |N:2  |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
    2 | DAKSHESH DARURI                 |6.0  |W  63|W  58|L   4|W  17|W  16|W  20|W   7|
   MI | 14598900 / R: 1553   ->1663     |N:2  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
    3 | ADITYA BAJAJ                    |6.0  |L   8|W  61|W  25|W  21|W  11|W  13|W  12|
   MI | 14959604 / R: 1384   ->1640     |N:2  |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
    4 | PATRICK H SCHILLING             |5.5  |W  23|D  28|W   2|W  26|D   5|W  19|D   1|
   MI | 12616049 / R: 1716   ->1744     |N:2  |W    |B    |W    |B    |W    |B    |B    |
-----------------------------------------------------------------------------------------
    5 | HANSHI ZUO                      |5.5  |W  45|W  37|D  12|D  13|D   4|W  14|W  17|
   MI | 14601533 / R: 1655   ->1690     |N:2  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
```

Given a player, the first row includes information about player's ID, Name, Total Points, oppoent ID's. The second row includes information about player's State, his/her Pre-ratings. Therefore, I decided to create two datasets - 1) **playerInfo1** from the first row and 2) **playerInfo2** from the second row.   

### playerInfo1 Table 
To extract player's ID, name, total points, opponents ID's, I read every 3rd line starting from row 5.
```{r }
playerInfo1 <-read.table(text = file[seq(5, length(file), by = 3)], header = FALSE, fill = TRUE, sep = "|")

kable(head(playerInfo1))
```

Each player can play up to 7 oppoents. V4 to V10 include the gamen results and oppoents ID's. the opponent ID's are obtained by extracting last two characters for each variable from V4 to V10 and then converting them to numeric variables.
```{r}
playerInfo1$V4 <- as.numeric(str_sub(playerInfo1$V4, -2, -1))
playerInfo1$V5 <- as.numeric(str_sub(playerInfo1$V5, -2, -1))
playerInfo1$V6 <- as.numeric(str_sub(playerInfo1$V6, -2, -1))
playerInfo1$V7 <- as.numeric(str_sub(playerInfo1$V7, -2, -1))
playerInfo1$V8 <- as.numeric(str_sub(playerInfo1$V8, -2, -1))
playerInfo1$V9 <- as.numeric(str_sub(playerInfo1$V9, -2, -1))
playerInfo1$V10 <- as.numeric(str_sub(playerInfo1$V10, -2, -1))
```

Then, I dropped V11, which is not needed and renamed headers for **playerInfo1**. Here are a couple rows from **playerInfo1**

```{r }
playerInfo1 <- subset(playerInfo1, select = -V11)
names(playerInfo1) <- c("playerID", "Name", "TotalPts", "Round1", "Round2", "Round3", "Round4", "Round5", "Round6", "Round7")

kable(head(playerInfo1))
```

#### Oppoent ID's
From playerInfo1, I subsetted playerID and round1 to round 7 to create oppontInfo data. Since the 7 possible oppenent ID's were saved in a wide format, I changed them in a long format using *gather()* function and created **oppoentInfo**. Tis way, the information about oppoents, in particular, the number of oppoents each player can be easily calculated. See the data structures for **opponentInfo** versus **opponentInfo_tidy**.
```{r}
opponentInfo <- subset(playerInfo1, select = c(playerID, Round1:Round7))
opponentInfo_tidy <- opponentInfo %>%
  gather(Round, opponentID, Round1:Round7) 

kable(head(opponentInfo))
kable(head(opponentInfo_tidy))
```

### playerInfo2
To extract player's state, pre-rating, I read every 3rd line starting from row 6 in the original text file.
```{r}
playerInfo2 <-read.table(text = file[seq(6, length(file), by = 3)], header = FALSE, fill = TRUE, sep = "|")
kable(head(playerInfo2))
```

Then, I created playerID and preRating that is extraced from V2 in **playerInfo2** In **playerInfo2** , I selected playerID, state, preRating and renamed the variables. See the final format of **playerInfo2**   

```{r}
playerInfo2$playerID <- seq(1, nrow(playerInfo2))
playerInfo2$preRating <- as.numeric(str_sub(playerInfo2$V2, 16, 19))
playerInfo2 <- data.frame(playerInfo2$playerID, playerInfo2$V1, playerInfo2$preRating)
names(playerInfo2) <- c("playerID", "State", "preRating")
kable(head(playerInfo2))
```

### Compute average ratings 
The average of opponents' pre-ratings can be computed by following two steps:
1. First, join two tables, **opponentInfo_tidy** and **playerInfo2** based on opponentID from **opponentInfo_tidy** and playerID from **playerInfo2**. This way, we will see the list of opponent ID's and their pre-ratings
2. Second, using the table from the first step, compute the average ratings by playerID

```{r}
opponentRatings <- sqldf('select t1.playerID, t1.opponentID, t2.preRating as opponentRating
                       from opponentInfo_tidy t1, playerInfo2 t2
                       where t1.opponentID = t2.playerID')

kable(head(opponentRatings))

averageRating <- sqldf('select playerID, round(avg(opponentRating)) as opponentAvgRating,
                        count(*) as numGames
                       from opponentRatings
                       group by playerID')
kable(head(averageRating))
kable(head(playerInfo1))
kable(head(playerInfo2))
```

### chessTournament and Create .csv file
**chessTournament** was finally created using three tables, **averageRating**, **playerInfo1** and **PlayerInfo2**. Then, it was exported as a .csv file. 

```{r}
chessTournament <- sqldf('select t1.playerID, t1.Name, t2.State, t1.TotalPts, 
                         t2.preRating, t3.opponentAvgRating
                         from playerInfo1 t1
                         inner join playerInfo2 t2
                         on t1.playerID = t2.playerID
                         inner join averageRating t3
                         on t1.playerID = t3.playerID')

kable(head(chessTournament))

write.csv(chessTournament, file = "H:/DataScience/CUNY/Spring2018/DATA607/project1/chessTournament.csv", row.names = FALSE)
```
