---
title: "Week13_Assignment"
author: "YoungKoung Kim"
date: "April 27, 2018"
output: html_document
---


```{r echo = FALSE, results='hide', message=FALSE, warning=FALSE}
mypw <- "XXXXX"
##install.packages("RMySQL")
##install.packages("htmlTable")
#install.packages("mongolite")

library(RMySQL)
library(htmlTable)
library(knitr)
library(mongolite)


```

### SQL to R dataframe 

Connect SQL flight database and create r data frame  

```{r warning=FALSE}
mydb = dbConnect(MySQL(), user = 'root', password = mypw,  dbname='flights', host='localhost')
#list of tables in the movies database
dbListTables(mydb)



airlines <- dbGetQuery(mydb, "SELECT * FROM airlines")
airports <- dbGetQuery(mydb,"SELECT * FROM airports;")
flights <- dbGetQuery(mydb,"SELECT * FROM flights;")
planes <- dbGetQuery(mydb,"SELECT * FROM planes;")
weather <-dbGetQuery(mydb,"SELECT * FROM weather;")

dbDisconnect(mydb) 

```

### MongoDB

#### Connect to MongoDB and Insert r dataframe to MongoDB


```{r}
# airlines
mongo_airlines  <- mongo(collection = "airlines", db= "flights", url = "mongodb://localhost")
mongo_airlines$drop()
mongo_airlines$insert(airlines)


# airports
mongo_airports  <- mongo(collection = "airports", db= "flights", url = "mongodb://localhost")
mongo_airports$drop()
mongo_airports$insert(airports)


# flights
mongo_flights  <- mongo(collection = "flights", db= "flights", url = "mongodb://localhost")
mongo_flights$drop()
mongo_flights$insert(flights)

# planes
mongo_planes  <- mongo(collection = "planes", db= "flights", url = "mongodb://localhost")
mongo_planes$drop()
mongo_planes$insert(planes)

# weather
mongo_weather  <- mongo(collection = "weather", db= "flights", url = "mongodb://localhost")
mongo_weather$drop()
mongo_weather$insert(weather)


```

#### Check the number of documents in MongoDB

```{r}
mongo_airlines$count()
mongo_airports$count()
mongo_flights$count()
mongo_planes$count()
mongo_weather$count()
```

#### Print out tables 
```{r}

kable(head(mongo_airlines$find()))
kable(head(mongo_airports$find()))
kable(head(mongo_flights$find()))
kable(head(mongo_planes$find()))
kable(head(mongo_weather$find()))
```

### Advantages and Disadvantages NoSQL database

#### Advantages
* NoSQL databases are non-relational and thus they are easier to manage and they provide a higher level of flexibility with newer data models.
* NoSQL databases are mostly Open Source and Low-Cost. They make rapid processing of real-time Big Data applications affordable.
* NoSQL databases have easier scalability through support for Map Reduce
* NoSQL databases do not need to develop a detailed database model. The non-relational nature of a NoSQL database allows database architects to quickly create a database without needing to develop a detailed (fine-grained) database model. This saves a lot of development time. 
Disadvantages

#### Disadvantages 
* While its continuing to grow rapidly, the NoSQL community is relatively new and lacks the maturity of the MySQL user base. 
* A major problem with NoSQL databases is the lack of reporting tools for analysis and performance testing. However, with MySQL, you can find a wide array of reporting tools to help you prove your application's validity. 
* NoSQL databases lack standardization and they need a standard query language like SQL. 

#### Reference
http://www.monitis.com/blog/cc-in-review-the-key-differences-between-sql-and-nosql-dbs/



