---
title: "Week3_Assignment"
author: "YoungKoung Kim"
date: "February 16, 2018"
output: html_document
---

### Problem 3. 
Copy the introductory example. The vector name stores the extracted names.

#### (a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.


```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```

##### Steps to create *first name*
1. Replace words that end with , or . 
2. Extract the words
```{r}
first_name <- str_replace(name, "(.*\\.)|(.*\\,)", "")
first_name <- unlist(str_extract(first_name, "\\w+"))
first_name
```
##### Steps to create *last name*
1. Extract last words or words that end with ',' 
2. Replace ',' with '' 
```{r}
last_name <- str_extract(name, "(\\w+($|,))")
last_name <- str_replace(last_name, "," , "")
last_name

first_last_name <- data.frame(first_name, last_name)
first_last_name
```

#### (b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.)

Directly match "Rev." or "Dr"

```{r}
writeLines(paste(str_detect(name, "(Rev.)|(Dr.)"), name, sep = ": \t\t"))
```
or Detect any strings with . that have at least 2 characters  
```{r}
writeLines(paste(str_detect(name, "([[:alpha:]]{2,}\\.)"), name, sep = ": \t\t"))
```

#### (c) Construct a logical vector indicating whether a character has a second name.
Second name starts with a blank and one character.

```{r}
writeLines(paste(str_detect(name, " [[:alpha:]]{1}\\."), name, sep = ": \t\t"))
```

### Problem 4. 
Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

#### (a) [0-9]+\\$
a charater with a digit followed by $ sign
```{r}
example_4a <- c("1900$ test", "some number$ error_Y2$K", "180$")
unlist(str_extract_all(example_4a, "[0-9]+\\$"))
```

#### (b) \\b[a-z]{1,4}\\b
a word that begins and ends with up to 4 lowercase letters       
```{r}
example_4b <-"R is free software and comes with ABSOLUTELY NO WARRANTY.You are welcome to redistribute it under certain conditions."
unlist(str_extract_all(example_4b, "\\b[a-z]{1,4}\\b"))
```

#### (c) .*?\\.txt$
any words that ends with ".txt"
```{r}
example_4c <-c("myfile.txt", "any text file", "week3_assignment.txt")
unlist(str_extract_all(example_4c, ".*?\\.txt$"))
```

#### (d) \\d{2}/\\d{2}/\\d{4}
two digits followed by '/', two digits followed by '/' and four digits. 
```{r}
example_4d <-c("2/16/2018", "02/16/2018", "00/00/0000", 'mm/dd/yyyy')
unlist(str_extract_all(example_4d, "\\d{2}/\\d{2}/\\d{4}"))
```

#### (e) <(.+?)>.+?</\\1>
strings that matches HTML tags
```{r}
example_4e <-c('<title>Automated Data Collection with R - Welcome</title>', "<html>Is this right?")
unlist(str_extract_all(example_4e, "<(.+?)>.+?</\\1>"))
```

### Problem 9. 
The following code hides a secret message. Crack it with R and regular expressions.

```{r}
raw_message  <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
secret_message <- unlist(str_extract_all(raw_message, "[[:upper:].! ]"))
secret_message <- str_replace_all(str_c(secret_message, collapse = ""), fixed("."), " ")
secret_message
```