---
title: "Week2_Assignment"
author: "YoungKoung Kim"
date: "February 10, 2018"
output: html_document
---

```{r echo = FALSE, results='hide', message=FALSE, warning=FALSE}
mypw <- "xxxxx"
library(RMySQL)
library(htmlTable)

```
### Introduction
When I saw this week's assingment, I was curious about movies that recently receieved high ratings by general movigoers. So, instead of asking friends/family/imaginary friends, I looked for a large dataset and found one from **movielens** (https://movielens.org/). **movielens** has very nice datasets that include movies and user ratings. Among their datasets, I chose the most recent data set, *'ml-latest-small'*. The dataset contains movie ratings by 671 users across 9125 movies between January 09, 1995 and October 16, 2016. A detailed decription of the data is available in http://files.grouplens.org/datasets/movielens/ml-latest-small-README.html

Note that I used the following packages for this assignment:
library(RMySQL)
library(htmlTable)

### Questions
Here are the questions that I would like to answer after analyzing the data:

  1. "What are the top 10 most popular movies?"
  2. "What is the movie that has the highest rating among these top 10 movies?"

### Database: movies
The database **movies** was created in SQL. In MySQL server, three datasets - 1) *movies*, 2) *ratings* and 3) *tags* - can be found in the **movies** database. 

```{r warning=FALSE}
mydb = dbConnect(MySQL(), user = 'root', password = mypw,  dbname='movies', host='localhost')
#list of tables in the movies database
dbListTables(mydb)

```

Now, let's take a look at each table in the **movies** database.
The first table is *movies*. It has three fields: movieID, title and genres. See the sample of a table with 5 records. 

```{r}
sampleMovies<-dbGetQuery(mydb, "select * from movies limit 5;")
htmlTable(sampleMovies)

```

The second table is *ratings*. It has four fields: userId, movieID, rating and timestamp. Rating is on a 5 star scale which ranges between 0.5 and 5.0. A user gives raings to multiple movies. See the sample of a table with 5 records. 
```{r warning=FALSE}
sampleRatings<-dbGetQuery(mydb, "select * from ratings limit 5;")
htmlTable(sampleRatings)

```

The third table is *tags*. It has three fields: userId, movieID, tag and timestamp. Tag is  a user-generated metadata about movies. See the sample of a table with 5 records. A user can provide multple tags to multiple movies. 
```{r}
sampleTags<-dbGetQuery(mydb, "select * from tags limit 5;")
htmlTable(sampleTags)

```
### Questions 1: "What are the top 10 most popular movies?"
To answer this question, in the table *movies*, I aggregated the table *ratings* to see how many users rated each movie and selected the 10 most popular movies.

First, I created a temporary table *popular_movie10*
```{r results='hide'}
#create temporary table popular_movie10 in mydb
dbSendQuery(mydb, "create temporary table popular_movie10
                  select movieId, count(userId) as num_users 
                  from ratings 
                  group by movieId
                  order by num_users desc
                  limit 10;")
```

Then, I got movie titles by joining *popular_movie10* with *movies*
```{r}

popular_movie10<-dbGetQuery(mydb, "select p.movieId, title, num_users, genres  
                                  from popular_movie10 p
                                  inner join movies m
                                  on p.movieId = m.movieId 
                                  order by num_users desc;")
htmlTable(popular_movie10)
```
So, **Forrest Gump (1994)** was the movie that had the largest number of users. 

### Questions 2: "What is the movie that has the highest rating among these top 10 movies?"
To answer this question, I created a temporary table *ratings_movie10* to obtain the average ratings for the top 10 most popular movies. 
```{r results='hide'}
#create temporary table ratings_movie10 in mydb
dbSendQuery(mydb, "create temporary table ratings_movie10
                   select m.movieId, title, avg(rating) as avg_rating
                   from ratings r
                   inner join movies m
                   on r.movieId = m.movieId
                   where r.movieId in (select movieId from popular_movie10)
                   group by m.movieId;")

```

Then, I got movie titles by joining *popular_movie10* with *ratings_movie10*
```{r warning=FALSE}

ratings_movie10<-dbGetQuery(mydb, "select p.movieId, title, num_users, avg_rating 
                                  from popular_movie10 p
                                  inner join ratings_movie10 r
                                  on p.movieId = r.movieId 
                                  order by avg_rating desc;")
htmlTable(ratings_movie10)
```
So, **Shawshank Redemption, The (1994)** had the highest rating among the top 10 most popular movies! 
